# Heterozygosity and Fis

The following code was used to analyze genetic diversity and inbreeding using the vcf file generated from the Filtering protocol. The steps used to calculate heterozygosity and Fis follow a [workflow developed by Jon Puritz](https://marineevoecolab.github.io/NACE_MAS_Genomics_Workshop/)\
\
```{r}
library(vcfR)
library(tidyverse)
library(reshape2)
library(dunn.test)
```


Reading in data from the vcf file created during filtering
```{r}
RISG <- read.vcfR("../Filtering/RISG.vcf")
```

We need a sample table for further analysis and to get a list of populations

```{r}
Samples <- read.table("Samples.txt", header = TRUE) %>% 
  mutate(Population = recode(Population, FI = 'CL1', NEH = 'CL2'))
poplist.names <- Samples$Population
```

## Calculating and Plotting Heterozygosity

To calculate heterozygosity we used the genetic_diff tool from vcfR

```{r}
het_results <- genetic_diff(RISG, pop=as.factor(poplist.names), method= 'nei')
round(colMeans(het_results[,c(3:8)], na.rm = TRUE), digits = 3)
```

To plot Heterozygosity we need to change the dataframe into one that can be plotted using ggplot

```{r}
het_df <- melt(het_results[,c(3:8)], varnames=c('Index', 'Sample'), value.name = 'Heterozygosity', na.rm=TRUE) %>%
  mutate(variable = recode(variable, Hs_CL1 = 'CL1', Hs_CL2 = 'CL2', Hs_MV = "MV", Hs_NR = "NR", Hs_GH = "GH", Hs_CT = "CT")) %>%
  mutate(Type = ifelse(variable == "CL1" | variable == "CL2", "Commercial", "Wild"))
```

Significance of heterozygoisty was calculated using a Kolmogorov-Smirnov test followed by a Dunn test. A Cohen's D test was also used to determine estimate of effect.
```{r}
#First we test for normality using a Kolmogorov-Smirnov Test
ks.test(het_df$Heterozygosity, "pnorm")
#The data is not normally distributed, we should proceed with a Kruskall test
kruskal.test(het_df$Heterozygosity ~ het_df$variable)
#The populations are significantly different, we should use a Dunn-Test post-hoc
dunn.test(het_df$Heterozygosity, g = het_df$variable, method  = "bonferroni")
#Checking estimate of effect with Cohen's D
cohens_d(het_df, Heterozygosity ~ variable)
```

Plotting Heterozygosity with ggplot, because the ranges of heterozygosity were narrow, the means and standard error were plotted along with median

```{r}
Het <- ggplot(het_df, aes(x=variable, y=Heterozygosity, fill = Type)) + geom_boxplot() +
  xlab("Population")+
  ylab("Observed Heterozygosity")+
  ggtitle("Observed Heterozygosity") + ylim(0,1)
plot(Het)
Hetsum <- het_df %>%
  group_by(variable) %>%
  summarize(avg = mean(Heterozygosity), se = sd(Heterozygosity)/sqrt(length(Heterozygosity)))
HetMean <- ggplot(Hetsum, aes(x=variable, y = avg))+ geom_point()+
  geom_errorbar(aes(ymin=avg-se, ymax=avg+se), width=.2, 
                position=position_dodge(0.05)) +
  xlab("Population")+
  ylab("Mean Heterozygosity")+
  ggtitle("Mean and Standard Error")+
  ylim(0,1)
plot(HetMean)
```

## Calculating and Plotting Fis

The --het function from Plink can be used to calculate F coefficient per individual. To calculate Fis, we need to use --het at a population level, which requires he plink file to be divided into multiple plink files by population. This can be done using a for loop and the --keep-fam funcion.
```{bash}
for pop in {FI,MV,CT,NR,NEH,GH}; do plink --file ../../Filtering/RISG.all.f --keep-fam ../../Filtering/$pop.samples --recode --out $pop.f; done
```

Next we do het for each population level plink file.
```{bash}
for pop in {FI,MV,CT,NR,NEH,GH}; do plink --file $pop.f --het --out $pop; done
```

Now we create a file that contains Fis for all individuals
```{bash}
head -1 CT.het > FIS.all
for pop in {FI,MV,CT,NR,NEH,GH}; do tail -n+2 $pop.het >> FIS.all; done
```

We then load this table into R
```{r}
library(tidyverse)
fis_df <- read_table("FIS/FIS.all") %>%
  mutate(FID = recode(FID, FI = "CL1", NEH = "CL2")) %>%
  mutate(Type = ifelse(FID == "CL1" | FID == "CL2", "Commercial", "Wild"))
```

Looking to see if any Fis values are statistically significant

```{r}
#First we test for normality using a Shapiro-Wilks Test
shapiro.test(fis_df$F)
#The data is not normally distributed, we should proceed with a Kruskall test
kruskal.test(fis_df$F ~ fis_df$FID)
#The populations are significantly different, we should use a Dunn-Test post-hoc
dunn.test(fis_df$F, g = fis_df$FID, method  = "bonferroni")
```

Plotting Fis

```{r}
Fissum <- fis_df %>%
  group_by(FID) %>%
  summarize(avg = mean(F), se = sd(F)/sqrt(length(F)), pop = n())
FIS <- ggplot(fis_df, aes(x=FID, y=F, fill = Type)) + geom_boxplot()+
  xlab("Population")+
  ylab("Inbreeding Coefficent (F)")+
  ggtitle("Inbreeding Coefficient per Population") + ylim(-1,1)
plot(FIS)
```


