## Percent of Polymorphic Loci

To determine the allele frequencies of each SNP in each population we can use the --frq function for the population level plink files generated in the Heterozygosity and Inbreeding protocol
```{bash}
for pop in {FI,MV,CT,NR,NEH,GH}; do plink --file $pop.f --frq --out $pop; done
```

To get the percent of polymorphic loci in each population, we take the number of SNPS in each population with a minor allele frequency greater than zero and divide that by 32,354 which is the number of SNPs kept after filtering. 
```{r}
library(tidyverse)
CT_ppl <- read.table("PPL/CT.frq", header = TRUE) %>%
  summarize(Population = "CT", Poly = sum(MAF != 0), Fixed = sum(MAF == 0), Po = sum(MAF != 0)/36968)%>%
  mutate(Type="Wild")
NR_ppl <- read.table("PPL/NR.frq", header = TRUE) %>%
  summarize(Population = "NR", Poly = sum(MAF != 0), Fixed = sum(MAF == 0), Po = sum(MAF != 0)/36968)%>%
  mutate(Type="Wild")
CL2_ppl <- read.table("PPL/NEH.frq", header = TRUE) %>%
  summarize(Population = "CL2", Poly = sum(MAF != 0), Fixed = sum(MAF == 0), Po = sum(MAF != 0)/36968)%>%
  mutate(Type="Commercial")
CL1_ppl <- read.table("PPL/FI.frq", header = TRUE) %>%
  summarize(Population = "CL1", Poly = sum(MAF != 0), Fixed = sum(MAF == 0), Po = sum(MAF != 0)/36968)%>%
  mutate(Type="Commercial")
MV_ppl <- read.table("PPL/MV.frq", header = TRUE) %>%
  summarize(Population = "MV", Poly = sum(MAF != 0), Fixed = sum(MAF == 0), Po = sum(MAF != 0)/36968)%>%
  mutate(Type="Wild")
GH_ppl <- read.table("PPL/GH.frq", header = TRUE) %>%
  summarize(Population = "GH", Poly = sum(MAF != 0), Fixed = sum(MAF == 0), Po = sum(MAF != 0)/36968)%>%
  mutate(Type="Wild")
All_ppl <- rbind(GH_ppl, NR_ppl, CT_ppl, CL1_ppl, CL2_ppl, MV_ppl)
```

Plotting percent of polymorphic loci
```{r}
Po_Plot <- ggplot(All_ppl, aes(x=Population, y=Po, fill = Type))+
 # geom_bar(stat="identity")+
  ggtitle("Proportion of Polymorphic Loci")+
  geom_bar_pattern(stat="identity", aes(fill = Type, pattern_alpha = Type), color = "black", lwd=1)+
  scale_pattern_alpha_manual(values = c(1, 0))+
  theme_bw()+
  scale_fill_grey()+
  theme(
    axis.ticks.length=unit(-0.2, "cm"),
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    legend.position = "none",
    panel.border = element_blank(),
    panel.grid.minor = element_blank(),text = element_text(face = "bold"),
    axis.line.x = element_line(size = 1, linetype = "solid", colour = "black"),
    axis.line.y = element_line(size = 1, linetype = "solid", colour = "black"))
plot(Po_Plot)
plot_grid(Po_Plot, Het, labels = c("A","B"))
```

